Efficiency

### 2018-01-08 DraftExec.Rnw

\begin{figure}
\includegraphics[width=\textwidth]{LaminarProp.png}
\caption{At very low speeds changes in regimes such as an increase in laminar flow over the blade section could lead to small changes in efficiency, perhaps explaining the distinct rise at low speeds for the properller at ~5kt in Figure \ref{fig:BMTEfficiency.png}}
\label{fig:LaminarProp.png}
\end{figure}

@@

\begin{figure}
<<plot_efficiency, echo=FALSE, warning=FALSE, message=F >>=
df <- suppressMessages(read_csv(find::this("/Submarines/eff-data-2018-01-03.csv")))

@@

\begin{figure}
<<plot_power1, echo=FALSE, warning=FALSE, message=F >>=
df <- read_rds("extdata/df_full_complete_fac.rds")
df %>%
  filter(var == "Power (kW)", pair %in% c("Central Selections")) %>%
  ggplot(aes(x = as.numeric(kts), y = as.numeric(val), col = cat))+
  scale_x_continuous(name ="speed (kt)")+
  scale_y_continuous(name = "Power (kW)")+
  labs(title = "Example Power Curves", col = "Power Demand")+
  geom_line(size = 0.8)
@
\caption{With central assumptions, a pumpjet powered submarine reaches 20kt with 6.5MW of delivered power.  Hotel load matches propulsion power demand around or below 5kt.}
\label{fig:power_curve}
\end{figure}

@@

\begin{figure}
<<plot_results1, echo=FALSE, warning=FALSE, message=F >>=
df <- read_rds("extdata/df_full_complete_fac.rds")
df %>%
  filter( pair %in% c("Widest Separation", "Narrowest Separation", "Central Selections")) %>%
  filter(var %in% c("Efficiency", "Endurance (hrs)", "Range (nm)")) %>%
  #filter(var == "end_prop", pair %in% c("top", "bottom", "propup", "propdown", "jetup", "jetdown", "wide", "narrow", "mid") ) %>%
  ggplot(aes(x = as.numeric(kts), y = as.numeric(val), linetype = as.factor(cat), col = level))+
  geom_line(aes(group = cat))+
  scale_linetype_manual(labels = c("Propeller", "Pumpjet" ), values = c(1, 2 ) )+
  scale_colour_manual(values = c("black", "blue", "red") )+
  scale_x_continuous(name ="speed (kt)")+
  scale_y_continuous(name =element_blank())+
  labs(linetype = "system")+
  facet_grid(var ~ pair, scales = "free", switch =  "y")
@
\caption{The most extreme and central assumptions are compared}
\label{fig:central_extreme}
\end{figure}

\begin{figure}
<<plot_results2, echo=FALSE, warning=FALSE, message=F >>=
df <- read_rds("extdata/df_full_complete_fac.rds")
df %>%
  filter( pair %in% c("Jet Lowered", "Jet Raised", "Central Selections")) %>%
  filter(var %in% c("Efficiency", "Endurance (hrs)", "Range (nm)")) %>%
  #filter(var == "end_prop", pair %in% c("top", "bottom", "propup", "propdown", "jetup", "jetdown", "wide", "narrow", "mid") ) %>%
  ggplot(aes(x = as.numeric(kts), y = as.numeric(val), linetype = as.factor(cat), col = level))+
  geom_line(aes(group = cat))+
  scale_linetype_manual(labels = c("Propeller", "Pumpjet" ), values = c(1, 2 ) )+
  scale_colour_manual(values = c("black", "blue", "red") )+
  scale_x_continuous(name ="speed (kt)")+
  scale_y_continuous(name =element_blank())+
  labs(linetype = "system")+
  facet_grid(var ~ pair, scales = "free", switch =  "y")
@
\caption{The impact of different jet efficiency assumptions compared to the central propeller curve shows wide variance}
\label{fig:central_jet}
\end{figure}

\begin{figure}
<<plot_results3, echo=FALSE, warning=FALSE, message=F >>=
df <- read_rds("extdata/df_full_complete_fac.rds")
df %>%
  filter( pair %in% c("Propeller Lowered", "Propeller Raised", "Central Selections")) %>%
  filter(var %in% c("Efficiency", "Endurance (hrs)", "Range (nm)")) %>%
  #filter(var == "end_prop", pair %in% c("top", "bottom", "propup", "propdown", "jetup", "jetdown", "wide", "narrow", "mid") ) %>%
  ggplot(aes(x = as.numeric(kts), y = as.numeric(val), linetype = as.factor(cat), col = level))+
  geom_line(aes(group = cat))+
  scale_linetype_manual(labels = c("Propeller", "Pumpjet" ), values = c(1, 2 ) )+
  scale_colour_manual(values = c("black", "blue", "red") )+
  scale_x_continuous(name ="speed (kt)")+
  scale_y_continuous(name =element_blank())+
  labs(linetype = "system")+
  facet_grid(var ~ pair, scales = "free", switch =  "y")
@
\caption{The impact of different propeller efficiency assumptions compared to the central pumpjet curve shows less variance}
\label{fig:central_prop}
\end{figure}

\begin{figure}
<<plot_results4, echo=FALSE, warning=FALSE, message=F >>=
df <- read_rds("extdata/df_full_complete_fac.rds")
df %>%
  filter( pair %in% c("Lowest Lines", "Highest Lines", "Central Selections")) %>%
  filter(var %in% c("Efficiency", "Endurance (hrs)", "Range (nm)")) %>%
  #filter(var == "end_prop", pair %in% c("top", "bottom", "propup", "propdown", "jetup", "jetdown", "wide", "narrow", "mid") ) %>%
  ggplot(aes(x = as.numeric(kts), y = as.numeric(val), linetype = as.factor(cat), col = level))+
  geom_line(aes(group = cat))+
  scale_linetype_manual(labels = c("Propeller", "Pumpjet" ), values = c(1, 2 ) )+
  scale_colour_manual(values = c("black", "blue", "red") )+
  scale_x_continuous(name ="speed (kt)")+
  scale_y_continuous(name =element_blank())+
  labs(linetype = "system")+
  facet_grid(var ~ pair, scales = "free", switch =  "y")
@
\caption{Comparing the lower and higher lines for both systems together}
\label{fig:bottom_top}
\end{figure}

@@

\begin{figure}
<<plot_proportions, echo=FALSE, warning=FALSE, message=F >>=
df <- read_rds("extdata/df_full_complete_fac.rds")
df %>%
  filter(var == "Advantage Propeller %", pair %in% c("Lowest Lines", "Jet Lowered", "Propeller Lowered", "Widest Separation", "Central Selections", "Narrowest Separation", "Propeller Raised", "Jet Raised", "Highest Lines") ) %>%
  ggplot(aes(x = as.numeric(kts), y = as.numeric(val), col = pair))+
  geom_line()+
  scale_x_continuous(name ="speed (kt)")+
  labs(col = "Curves Compared")+
  scale_y_continuous(name ="Percentage Propeller Advantage", breaks=seq(0, 120, 10))
@
\caption{The most proportional differences between the propeller and pumpjet performances for a range of assumptions}
\label{fig:all_proportions}
\end{figure}

@@

\subsection{Sensitivity to Hotel Load}
As can be seen in Figure \ref{fig:sens_hotel} the difference between the two systems is very sensitive to changes in hotel load, both in proportional and nominal terms.  The impact of halving the hotel load from 100kw to 50kw more than doubles the \textit{difference} between a pumpjet and propeller dived endurances, from around two days to well over four days at speeds between 2.5 and 3.5kt.  In the context of a submarine which might have a nominal dived requirement at patrol speeds in the order of several days, this suggests that extremely large gains can be made by reducing hotel load.  Increasing the hotel load has the noticeable effect of pushing the speed of largest difference to higher speeds, and substantially reducing the total difference. If the hotel load was doubled to 200kW, the middle efficiency curve pair would still see a difference of nearly a day between the two systems, or around 19.5 hours at speeds of 4.5-5 kts.

@@

It's important to acknowledge that such substantial differences in hotel load merit serious consideration, both in terms system design, as well as for variance in operation of a submarine with a given system design.  As has been discussed, the precise hotel load of submarines is not generally publically available, which justifies senitivity testing around a relatively wide potential range.  The size of the crew required for different designs might vary considerably, as it also would for different missions (for example with the embarkation of a special forces congingent) which would necessarily have considerable impact on energy demand for life support, including sanitation, CO2 scrubbing, and cooking. In addition, it is broadly understood that the AN/BYG-1 combat system consumes substantially more power than other alternatives, reportedly in the order of 100kW \parencite{patrick2011}.

Furthermore, the hotel load may be varied considerably at the captain's discretion depending on operational requirements.  Not all sensors or components of the combat system need to run at all times, for example during long transits in order to reach an operational area, or during a long dive in order to escape an operational area once a large strike has already been conducted.  In wartime, when surfacing to snort within range of enemy airpower or close to naval bases might result in the sumbarine probably being sunk, the crew might willingly forego hot showers, hot meals, and climate-controlled air in order to increase their chances of survival.  Consequently, it is perfectly plausible that changes of hotel load in the order of a factor of two for tactically significant time periods (days) is achievable.  The results of this sensitivity analysis show that the penalty inflicted by the pumpjet will be substantially greater in such circumstances.

The proportional impact of the changes in hotel load area also substantial, and exhibit clearly the same trend of the impact being larger at higher speeds for higher hotel loads. For the efficiency curve pairs that comprise the central selections, decreasing the hotel load to 50kt would see the dived endurance and range of the submarine improved by the adoption of propellers 80\% between 4 and 5.5kt. A much higher hotel load, at 200kW, would see that peak difference over 40\% between 5.5 and 8kt.

It should also be noted that sensitivity testing has not been shown in this paper around alternative values of the drag coefficient $C_d$ which was solved for by matching peak power output and top speed. It is considered relatively unlikely that such dramatic changes in this constant are plausible.  Furthermore, changing this constant will have a very similar impact to alerations in hotel load, in terms of its affect on the speed where greatest differences will occur and their probable magnitude.  Consequently, consideration of the sensitivity to hotel load is likely to easily encompass the plausible extent, and nature, of the likely uncertainty in this factor.

\begin{figure}
<<plot_sens_hotel, echo=FALSE, warning=FALSE, message=F >>=
df <- read_rds("extdata/df_full_diff_cornered.rds")
df %>%
  filter(test == "hotel") %>%
  ggplot(aes(x = kts, y = val, col = pair))+
  geom_line()+
  labs(col = "Efficiency \nCurve Pair")+
  scale_y_continuous(name = element_blank())+
  facet_grid(var ~ as.factor(hotel), scales = "free")
@
\caption{The impact of the propulsion system efficiencies are very sensitive to changes in the hotel load, including as a proportion.}
\label{fig:sens_hotel}
\end{figure}

@@

@@

@@

\begin{figure}
<<plot_sens_battery, echo=FALSE, warning=FALSE, message=F >>=
df <- read_rds("extdata/df_full_diff_cornered.rds")
df %>%
  filter(test == "battery") %>%
  ggplot(aes(x = kts, y = val, col = pair))+
  geom_line()+
  labs(col = "Efficiency \nCurve Pair")+
  scale_y_continuous(name = element_blank())+
  facet_grid(var ~ as.factor(battery), scales = "free")
@
\caption{The impact of the propulsion system efficiencies are extremely sensitive to battery energy density (or load) though proportions remain fixed.}
\label{fig:sens_battery}
\end{figure}

@@

\begin{figure}
<<plot_sens_corners, echo=FALSE, warning=FALSE, message=F >>=
df <- read_rds("extdata/df_full_diff_cornered.rds")
df %>%
  filter(test == "corners") %>%
  ggplot(aes(x = kts, y = val, col = pair))+
  geom_line()+
  labs(col = "Efficiency \nCurve Pair")+
  scale_y_continuous(name = element_blank())+
  theme(legend.position="bottom")+
  facet_grid(var ~ corner, scales = "free", switch = "y")
@
\caption{Specifying low density batteries and a high hotel load dramatically reduces the proportional and nominal impact of propulsion system efficiencies at low speeds}
\label{fig:sens_corners}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\textwidth]{InceptionDepth.png}
\caption{Increasing in depth improves the speed achievable before cavitation begins to occur, as shown in the case of a torpedo \parencite{suryanarayana2010cavitation}}
\label{fig:InceptionDepth.png}
\end{figure}

@@

\begin{figure}
<<plot_inception, echo=FALSE, warning=F, message=F>>=
speed <- rep(seq(0.5,20,0.5),90)
VP. <- c(rep(1.228*1000,40), rep(2.339*1000,40), rep(4.245*1000,40))
VP <- rep(VP., 30)
temp. <- c(rep(10,40), rep(20,40), rep(30,40))
temp <- rep(temp., 30)
depth <- c(rep(10,120), rep(20,120), rep(30,120), rep(40,120), rep(50,120), rep(60,120), rep(70,120), rep(80,120), rep(90,120), rep(100,120),
           rep(110,120), rep(120,120), rep(130,120), rep(140,120), rep(150,120), rep(160,120), rep(170,120), rep(180,120), rep(190,120), rep(200,120),
           rep(210,120), rep(220,120), rep(230,120), rep(240,120), rep(250,120), rep(260,120), rep(270,120), rep(280,120), rep(290,120), rep(300,120))

@@

\begin{figure}
\includegraphics[width=\textwidth]{VortexShedding.png}
\caption{Vortex shedding following a flow separation leads to eddies being released unsteadily into the main flow stream, as shown here on the exterior of an accelerating duct.  In the case of decelerating duct (pumpjet), duct shape is inverted, leading to the vortices being shed into the impeller blades. \parencite{willemsen2013}}
\label{fig:VortexShedding.png}
\end{figure}

@@

\subsection{The impact of duct shielding}
The presence of the duct around the impeller may also have the effect of preventing the noise from being effectively radiated away from the propulsor.  This may well result in substantial reductions in some types of noise being radiated away from the propulsor in certain directions, which may have given rise to the consistent qualification of 'radiated' being used to descrive the type of noise which differentiated the French submarine bid \parencite{stewart2016} \parencite{davies2017} \parencite{ohff2016}, including by the designers of the submarine \parencite{autret2016}.  This is a secondary effect, quite separate to the avoidance of cavitation, which might be seen as an acoustic advantage in circumstances where cavitation might not be expected to occur on either type of propulsor.  It is highly likely that this is the effect which is relied upon to sustain claims that a pumpjet is generally acoustically superior to a propeller, including at low speeds.

>>>>

\begin{figure}[h]
\captionsetup{width=0.4\textwidth}
\includegraphics[width=0.4\textwidth, center]{LaminarProp.png}
\caption{At very low speeds changes in regimes such as an increase in laminar flow over the blade section could lead to small changes in efficiency, perhaps explaining the distinct rise at low speeds for the properller at ~5kt in Figure \ref{fig:BMTEfficiency.png}}
\label{fig:LaminarProp.png}
\end{figure}

@@

\begin{figure}[p]
<<plot_efficiency, echo=FALSE, warning=FALSE, message=F >>=
df <- suppressMessages(read_csv(find::this("/Submarines/eff-data-2018-01-03.csv")))

@@

\begin{figure}[p]
<<plot_power1, echo=FALSE, warning=FALSE, message=F >>=
df <- read_rds("extdata/df_full_complete_fac.rds")
df %>%
  filter(var == "Power (kW)", pair %in% c("Central Selections")) %>%
  ggplot(aes(x = as.numeric(kts), y = as.numeric(val), col = cat))+
  scale_x_continuous(name ="speed (kt)")+
  scale_y_continuous(name = "Power (kW)")+
  labs(title = "Example Power Curves", col = "Power Demand")+
  geom_line(size = 0.8)
@
\caption{With central assumptions, a pumpjet powered submarine reaches 20kt with 6.5MW of delivered power.  Hotel load matches propulsion power demand around or below 5kt.}
\label{fig:power_curve}
\end{figure}

@@

\begin{figure}[p]
<<plot_results1, echo=FALSE, warning=FALSE, message=F >>=
df <- read_rds("extdata/df_full_complete_fac.rds")
df %>%
  filter( pair %in% c("Widest Separation", "Narrowest Separation", "Central Selections")) %>%
  filter(var %in% c("Efficiency", "Endurance (hrs)", "Range (nm)")) %>%
  #filter(var == "end_prop", pair %in% c("top", "bottom", "propup", "propdown", "jetup", "jetdown", "wide", "narrow", "mid") ) %>%
  ggplot(aes(x = as.numeric(kts), y = as.numeric(val), linetype = as.factor(cat), col = level))+
  geom_line(aes(group = cat))+
  scale_linetype_manual(labels = c("Propeller", "Pumpjet" ), values = c(1, 2 ) )+
  scale_colour_manual(values = c("black", "blue", "red") )+
  scale_x_continuous(name ="speed (kt)")+
  scale_y_continuous(name =element_blank())+
  labs(linetype = "system")+
  facet_grid(var ~ pair, scales = "free", switch =  "y")
@
\caption{The most extreme and central assumptions are compared}
\label{fig:central_extreme}
\end{figure}

\begin{figure}[p]
<<plot_results2, echo=FALSE, warning=FALSE, message=F >>=
df <- read_rds("extdata/df_full_complete_fac.rds")
df %>%
  filter( pair %in% c("Jet Lowered", "Jet Raised", "Central Selections")) %>%
  filter(var %in% c("Efficiency", "Endurance (hrs)", "Range (nm)")) %>%
  #filter(var == "end_prop", pair %in% c("top", "bottom", "propup", "propdown", "jetup", "jetdown", "wide", "narrow", "mid") ) %>%
  ggplot(aes(x = as.numeric(kts), y = as.numeric(val), linetype = as.factor(cat), col = level))+
  geom_line(aes(group = cat))+
  scale_linetype_manual(labels = c("Propeller", "Pumpjet" ), values = c(1, 2 ) )+
  scale_colour_manual(values = c("black", "blue", "red") )+
  scale_x_continuous(name ="speed (kt)")+
  scale_y_continuous(name =element_blank())+
  labs(linetype = "system")+
  facet_grid(var ~ pair, scales = "free", switch =  "y")
@
\caption{The impact of different jet efficiency assumptions compared to the central propeller curve shows wide variance}
\label{fig:central_jet}
\end{figure}

\begin{figure}[p]
<<plot_results3, echo=FALSE, warning=FALSE, message=F >>=
df <- read_rds("extdata/df_full_complete_fac.rds")
df %>%
  filter( pair %in% c("Propeller Lowered", "Propeller Raised", "Central Selections")) %>%
  filter(var %in% c("Efficiency", "Endurance (hrs)", "Range (nm)")) %>%
  #filter(var == "end_prop", pair %in% c("top", "bottom", "propup", "propdown", "jetup", "jetdown", "wide", "narrow", "mid") ) %>%
  ggplot(aes(x = as.numeric(kts), y = as.numeric(val), linetype = as.factor(cat), col = level))+
  geom_line(aes(group = cat))+
  scale_linetype_manual(labels = c("Propeller", "Pumpjet" ), values = c(1, 2 ) )+
  scale_colour_manual(values = c("black", "blue", "red") )+
  scale_x_continuous(name ="speed (kt)")+
  scale_y_continuous(name =element_blank())+
  labs(linetype = "system")+
  facet_grid(var ~ pair, scales = "free", switch =  "y")
@
\caption{The impact of different propeller efficiency assumptions compared to the central pumpjet curve shows less variance}
\label{fig:central_prop}
\end{figure}

\begin{figure}[p]
<<plot_results4, echo=FALSE, warning=FALSE, message=F >>=
df <- read_rds("extdata/df_full_complete_fac.rds")
df %>%
  filter( pair %in% c("Lowest Lines", "Highest Lines", "Central Selections")) %>%
  filter(var %in% c("Efficiency", "Endurance (hrs)", "Range (nm)")) %>%
  #filter(var == "end_prop", pair %in% c("top", "bottom", "propup", "propdown", "jetup", "jetdown", "wide", "narrow", "mid") ) %>%
  ggplot(aes(x = as.numeric(kts), y = as.numeric(val), linetype = as.factor(cat), col = level))+
  geom_line(aes(group = cat))+
  scale_linetype_manual(labels = c("Propeller", "Pumpjet" ), values = c(1, 2 ) )+
  scale_colour_manual(values = c("black", "blue", "red") )+
  scale_x_continuous(name ="speed (kt)")+
  scale_y_continuous(name =element_blank())+
  labs(linetype = "system")+
  facet_grid(var ~ pair, scales = "free", switch =  "y")
@
\caption{Comparing the lower and higher lines for both systems together}
\label{fig:bottom_top}
\end{figure}

@@

\begin{figure}[p]
<<plot_proportions, echo=FALSE, warning=FALSE, message=F >>=
df <- read_rds("extdata/df_full_complete_fac.rds")
df %>%
  filter(var == "Advantage Propeller %", pair %in% c("Lowest Lines", "Jet Lowered", "Propeller Lowered", "Widest Separation", "Central Selections", "Narrowest Separation", "Propeller Raised", "Jet Raised", "Highest Lines") ) %>%
  ggplot(aes(x = as.numeric(kts), y = as.numeric(val), col = pair))+
  geom_line()+
  scale_x_continuous(name ="speed (kt)")+
  labs(col = "Curves Compared")+
  scale_y_continuous(name ="Percentage Propeller Advantage", breaks=seq(0, 120, 10))
@
\caption{The most proportional differences between the propeller and pumpjet performances for a range of assumptions}
\label{fig:all_proportions}
\end{figure}

@@

\subsection{Sensitivity to Hotel Load}

As can be seen in Figure \ref{fig:sens_hotel} the difference between the two systems is very sensitive to changes in hotel load, both in proportional and nominal terms.  The impact of halving the hotel load from 100kw to 50kw more than doubles the \textit{difference} between a pumpjet and propeller dived endurances, from around two days to well over four days at speeds between 2.5 and 3.5kt.  In the context of a submarine which might have a nominal dived requirement at patrol speeds in the order of several days, this suggests that extremely large gains can be made by reducing hotel load.  Increasing the hotel load has the noticeable effect of pushing the speed of largest difference to higher speeds, and substantially reducing the total difference. If the hotel load was doubled to 200kW, the middle efficiency curve pair would still see a difference of nearly a day between the two systems, or around 19.5 hours at speeds of 4.5-5 kts.

\begin{figure}[p]
<<plot_sens_hotel, echo=FALSE, warning=FALSE, message=F >>=
df <- read_rds("extdata/df_full_diff_cornered.rds")
df %>%
  filter(test == "hotel") %>%
  ggplot(aes(x = kts, y = val, col = pair))+
  geom_line()+
  labs(col = "Efficiency \nCurve Pair")+
  scale_y_continuous(name = element_blank())+
  facet_grid(var ~ as.factor(hotel), scales = "free")
@
\caption{The impact of the propulsion system efficiencies are very sensitive to changes in the hotel load, including as a proportion.}
\label{fig:sens_hotel}
\end{figure}

@@

@@

Changes in the effective battery energy density will have the same effect as having larger masses of battery stored.  They simply directly impact the total amount of energy available for consumption before recharging batteries while snorkeling. There is therefore no shift in the proportional impact on range and endurance due to this factor, as can be seen in the lower graphs in Figure \ref{fig:sens_battery}.  The nominal impact, however increases directly and linearly with the total energy stored.

\begin{figure}[p]
<<plot_sens_battery, echo=FALSE, warning=FALSE, message=F >>=
df <- read_rds("extdata/df_full_diff_cornered.rds")
df %>%
  filter(test == "battery") %>%
  ggplot(aes(x = kts, y = val, col = pair))+
  geom_line()+
  labs(col = "Efficiency \nCurve Pair")+
  scale_y_continuous(name = element_blank())+
  facet_grid(var ~ as.factor(battery), scales = "free")
@
\caption{The impact of the propulsion system efficiencies are extremely sensitive to battery energy density (or load) though proportions remain fixed.}
\label{fig:sens_battery}
\end{figure}

The results in nominal terms are striking in their magnitude.  In terms of enduracy, doubling the stored electrical energy, for example by adopting lithium batteries, increase the \textit{difference} between the two systems under the central efficiency curve assumptions to nearly four days (93 hours) at 4kt.  Given that most modern, capable submarines are estimated to only be able to achieve dived endurances between 60 and 100 hours on batteries at 4kt \parencite{buckingham2008submarine} this cannot be dismissed in terms of its tactical significance.  Even under the eficiency curve assumptions representing the narrowest spearation considered, the difference in this case would be 59 hours, almost equal to the estimated dived endurance of a Type214 operating without AIP.

The in terms of submerged range, the same proportional impact is also extremely significant.  Doubling the energy stored inceases the \textit{difference} between two otherwise identical submarines to over 400nm under central assumptions at 4.5kt.  That's a larger distance than the width of the Java Sea, the most central body of water in the Indonesian Archipelago, or most of the length of the Malacca Strait.  The tactical significance of such distances are not likely to be lightly dismissed. If far smaller amounts of energy are embarked, the nominal differences reduce to perhaps 100nm, in direct proportion to the reduction of the submarine's dived endurance.  This is still a non-trivial distance in terms of it's tactical signicance, and comprises a substantial fraction of the dived ranges of most non-AIP submarines.  It still represents a significant difference in operational terms, equal approximately to the width of Hainan Island, or neraly twice the length of the Sunda Strait.

@@

Only brief is given here to the impact of Air Independent Propulsion, which has substantially higher energy densities possible even than Lithium, and also amounts to the addtional embarkation of greater mass. The volumetric energy density of the feedstocks to modern AIP systems (liquid oxygen and methanol) is estimated to be approximately ten times greater than that of lead-acid batteries \parencite{davies2016}, around 4.5MJ/L.  Given the greater mass embarked, and additional studies suggesting that at constant speeds they might extend the dived endurance of a modern submarine from around 60 hours to over 1000 hours (a factor of roughly sixteen) an approximate estimation of the the direction and magnitude of this impact may by scaling up energy density by this factor.  This would amount to adding three further frames to the right of Figure \ref{fig:sens_battery}, with the energy density being doubled in each successive frame. This would plausibly (under the central line selections) amount to more than month of dived endurance \textit{difference} between AIP inclusive submarines with propellers or pumpjets, amounting to thousands of nautical miles of continuously submerged travel.  This difference would also now be material in terms of the overall range of the submarine amounting to a further 15-20\% reduction in the submarine's overall range.

@@

@@

Writing a paragraph here...

\begin{figure}[p]
<<plot_sens_corners, echo=FALSE, warning=FALSE, message=F>>=
df <- read_rds("extdata/df_full_diff_cornered.rds")
df %>%
  mutate(corner = str_replace(corner, ", ", "\n")) %>%
  filter(test == "corners") %>%
  ggplot(aes(x = kts, y = val, col = pair))+
  geom_line()+
  labs(col = "Efficiency \nCurve Pair")+
  scale_y_continuous(name = element_blank())+
  theme(legend.position="bottom")+
  facet_grid(var ~ corner, scales = "free", switch = "y")
@
\caption{Specifying low density batteries and a high hotel load dramatically reduces the proportional and nominal impact of propulsion system efficiencies at low speeds}
\label{fig:sens_corners}
\end{figure}

@@

\begin{figure}[h]
\captionsetup{width=0.7\textwidth}
\includegraphics[width=0.7\textwidth, center]{InceptionDepth.png}
\caption{Increasing in depth improves the speed achievable before cavitation begins to occur, as shown in the case of a torpedo \parencite{suryanarayana2010cavitation}}
\label{fig:InceptionDepth.png}
\end{figure}

@@

\begin{figure}[p]
<<plot_inception, echo=FALSE, warning=F, message=F>>=
speed <- rep(seq(0.5,20,0.5),90)
VP. <- c(rep(1.228*1000,40), rep(2.339*1000,40), rep(4.245*1000,40))
VP <- rep(VP., 30)
temp. <- c(rep(10,40), rep(20,40), rep(30,40))
temp <- rep(temp., 30)
depth <- c(rep(10,120), rep(20,120), rep(30,120), rep(40,120), rep(50,120), rep(60,120), rep(70,120), rep(80,120), rep(90,120), rep(100,120),
           rep(110,120), rep(120,120), rep(130,120), rep(140,120), rep(150,120), rep(160,120), rep(170,120), rep(180,120), rep(190,120), rep(200,120),
           rep(210,120), rep(220,120), rep(230,120), rep(240,120), rep(250,120), rep(260,120), rep(270,120), rep(280,120), rep(290,120), rep(300,120))

@@

\begin{figure}[h]
\captionsetup{width=0.5\textwidth}
\includegraphics[width=0.6\textwidth, center]{VortexShedding.png}
\caption{Vortex shedding following a flow separation leads to eddies being released unsteadily into the main flow stream, as shown here on the exterior of an accelerating duct.  In the case of decelerating duct (pumpjet), duct shape is inverted, leading to the vortices being shed into the impeller blades. \parencite{willemsen2013}}
\label{fig:VortexShedding.png}
\end{figure}

@@

\subsection{The impact of duct shielding}
The presence of the duct around the impeller may also have the effect of preventing the noise from being effectively radiated away from the propulsor.  This may well result in substantial reductions in some types of noise being radiated away from the propulsor in certain directions, which may have given rise to the consistent qualification of 'radiated' being used to descrive the type of noise which differentiated the French submarine bid (\cite{stewart2016}, \cite{davies2017}, \cite{ohff2016}), including by the designers of the submarine \parencite{autret2016}.  This is a secondary effect, quite separate to the avoidance of cavitation, which might be seen as an acoustic advantage in circumstances where cavitation might not be expected to occur on either type of propulsor.  It is highly likely that this is the effect which is relied upon to sustain claims that a pumpjet is generally acoustically superior to a propeller, including at low speeds.
